HO unification from object language to meta language

### 2024-02-27 paper/main.tex
wip

\noindent
The \elpiIn{name} builtin predicate tests if a term is a bound variable. The
compiler~\ref{sec:xxx} needs to support terms outside \llambda for practical
reasons, so we don't assume all out terms are in \llambda but rather test.

@@

\begin{elpicode}
typeabbrev memory A (list (option A)).
type set?   nat -> memory A -> A -> o.
type unset  nat -> memory A -> o.
type assign nat -> memory A -> A -> memory A -> o.
\end{elpicode}

@@

\begin{elpicode}
type fderef fsubst -> fm -> fm -> o.
% flattening
fderef S (fapp [fuva N|A]) R :- set? N S (fapp B), !,
  append A B C, fderef S (fapp C) R.
% traversal
fderef S (fapp A) (fapp B) :- map (fderef S) A B.
fderef S (flam F) (flam G) :- 
  pi x\ fderef S x x => fderef S (F x) (G x).
fderef _ (fcon C) (fcon C).
% dereferencing
fderef S (fuva N) R :- set? N S T, fderef S T R.
fderef _ (fuva N) (fuva N) :- unset N S.
\end{elpicode}

@@

\begin{elpicode}
type deref subst -> tm -> tm -> o.
% ... similar to above ...
deref S (uva N A) R :- set? N S F, move F A T, deref S T R.
deref S (uva N A) (uva N B) :- unset N S, map (deref S) A B.
type move assignment -> list tm -> tm -> o.
move (abs F) [X | L] R :- move (F X) L R.
move (val A)         [] A :- !.
move (val (app A))   L (app X) :- append A L X.
move (val (uva N A)) L (uva N X) :- append A L X.
move (val (con H))   L (app [con H | L]).
\end{elpicode}

@@

\begin{figure*}
\begin{elpicode}
~\PYG{k+kd}{type} \PYG{n+nf}{(\Ue)} \PYG{k+kt}{tm -> tm -> subst -> subst -> o}~.
% Congruence
(app A ~\Ue~app B) S S1 :- fold2 ~(\Ue)~A B S S1.
(lam F ~\Ue~lam G) S S1 :- pi x\ (pi S\ (x ~\Ue~x) S S) => (F x ~\Ue~G x) S S1.
~\Ue~(con C) (con C) S S.
% deref
(uva N A ~\Ue~T) S S1 :- set? N S F, move F A T1, (T1 ~\Ue~T) S S1.
(T ~\Ue~uva N A) S S1 :- set? N S F, move F A T1, (T ~\Ue~T1) S S1.
% flex-flex
(uva N A ~\Ue~uva M B) S S3 :-  unset M, unset N,
  distinct-names A, distinct-names B, 
  new S W S1, prune W Args1 B Ass, 
  assign N S1 Ass S2, assign M S2 Ass S3.
% assignment
(uva N A ~\Ue~T) S S1 :- distinct-names A, not (T = uva _ _), not_occ N S T,  
  bind A T T1, assign N S T1 S1.
(T ~\Ue~uva N A) S S1 :- distinct-names A, not (T = uva _ _), not_occ N S T, 
  bind A T T1, assign N S T1 S1.

@@

typeabbrev memory A (list (option A)).
type set? nat -> memory A -> A -> o.
set? N S T :- nth N S (some T).
type unset nat -> memory A -> o.
unset N S :- nth N S none.
type assign nat -> memory A -> A -> memory A -> o.
assign z [none|M] T [some T|M].
assign (s N) [X|M] T [X|M1] :- assign N M T M1. 
kind nat type.
type z nat.
type s nat -> nat.
type nth nat -> list A -> A -> o.
nth z [X|_] X.
nth (s N) [_|L] X :- nth N L X.

type new memory A -> nat -> memory A -> o.
type prune .
type move .
type beta.
TODO

>>>>

\noindent
The \elpiIn{name} builtin predicate tests if a term is a bound variable.
\footnote{one could always load name x for every x under a pi and get rid of the name builtin}
The
compiler~\ref{sec:xxx} needs to support terms outside \llambda for practical
reasons, so we don't assume all out terms are in \llambda but rather test.

@@

\begin{elpicode}
typeabbrev memory A (list (option A)).
type set?   nat -> memory A -> A -> o.
type unset? nat -> memory A -> o.
type assign nat -> memory A -> A -> memory A -> o.
\end{elpicode}

@@

\begin{elpicode}
type fderef fsubst -> fm -> fm -> o.
% flattening
fderef S (fapp [fuva N|A]) R :- set? N S (fapp B), !,
  append A B C, fderef S (fapp C) R.
% traversal
fderef S (fapp A) (fapp B) :- map (fderef S) A B.
fderef S (flam F) (flam G) :- 
  pi x\ fderef S x x => fderef S (F x) (G x).
fderef _ (fcon C) (fcon C).
% dereferencing
fderef S (fuva N) R :- set? N S T, fderef S T R.
fderef S (fuva N) (fuva N) :- unset? N S.
\end{elpicode}

@@

\begin{elpicode}
type deref subst -> tm -> tm -> o.
% ... similar to above ...
deref S (uva N A) R :- set? N S F, move F A T, deref S T R.
deref S (uva N A) (uva N B) :- unset? N S, map (deref S) A B.
type move assignment -> list tm -> tm -> o.
move (abs F) [X | L] R :- move (F X) L R.
move (val A)         [] A :- !.
move (val (app A))   L (app X) :- append A L X.
move (val (uva N A)) L (uva N X) :- append A L X.
move (val (con H))   L (app [con H | L]).
\end{elpicode}

@@

\begin{figure*}
\begin{elpicode}
~\PYG{k+kd}{type} \PYG{n+nf}{(\Ue)} \PYG{k+kt}{tm -> tm -> subst -> subst -> o}~.
% Congruence
(app A ~\Ue~app B) S S1 :- fold2 ~(\Ue)~A B S S1.
(lam F ~\Ue~lam G) S S1 :- pi x\ (pi S\ (x ~\Ue~x) S S) => (F x ~\Ue~G x) S S1.
~\Ue~(con C) (con C) S S.
% deref
(uva N A ~\Ue~T) S S1 :- set? N S F, move F A T1, (T1 ~\Ue~T) S S1.
(T ~\Ue~uva N A) S S1 :- set? N S F, move F A T1, (T ~\Ue~T1) S S1.
% flex-flex
(uva N A ~\Ue~uva M B) S S3 :-  unset? M, unset? N,
  distinct-names A, distinct-names B, 
  new S W S1, prune W Args1 B Ass, 
  assign N S1 Ass S2, assign M S2 Ass S3.
% assignment
(uva N A ~\Ue~T) S S1 :- distinct-names A, not (T = uva _ _), not_occ N S T,  
  bind A T T1, assign N S T1 S1.
(T ~\Ue~uva N A) S S1 :- distinct-names A, not (T = uva _ _), not_occ N S T, 
  bind A T T1, assign N S T1 S1.

@@

typeabbrev memory A (list (option A)).
type set? nat -> memory A -> A -> o.
set? N S T :- nth N S (some T).
type unset? nat -> memory A -> o.
unset? N S :- nth N S none.
type assign nat -> memory A -> A -> memory A -> o.
assign z [none|M] T [some T|M].
assign (s N) [X|M] T [X|M1] :- assign N M T M1. 
kind nat type.
type z nat.
type s nat -> nat.
type nth nat -> list A -> A -> o.
nth z [X|_] X.
nth (s N) [_|L] X :- nth N L X.

type new memory A -> nat -> memory A -> o.
type prune .
type move .
type beta.
type not_occ.
TODO
